---
title: "Ang Akkermansia ASVs"
author: "Katherine Mueller"
date: "7/8/2021"
output: html_document
---

#Load Libraries
```{r}
library(phyloseq)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

#Define Paths
```{r}
ang.rds <- "../../Data/Analysis/ang.rds"
out.dir <- "../../Data/Anslysis"
```

#Load and Check Phyloseq Object
```{r}
ang.ps = read_rds(ang.rds)
print(ang.ps)
```

```{r}
sample_variables(ang.ps)
```

#Relative Abundance
```{r}
rel.ps <- ang.ps %>%
  transform_sample_counts(function(x) x/sum(x))
```


#Akkermansia ASV Identification
```{r}
akks.ps <- rel.ps %>%
  subset_taxa(Genus == "Akkermansia")
print(akks.ps)
```

Filter for Akk ASVs at a mean abundance of at least 0.01%
```{r}
mean_cutoff = 0.0001
mean_test = function(x) {
  mean(x) >= mean_cutoff
}

akks.0001mean.ps <- akks.ps %>%
  filter_taxa(mean_test, prune = TRUE)
print(akks.0001mean.ps)
```

```{r}
akks0001 <- akks.0001mean.ps %>%
  otu_table() %>%
  as.data.frame()
colnames(akks0001)
```

Double check that none of the other ASVs should be included
```{r}
#Get dataframe for akk ASV abundances
all_akks <- akks.ps %>%
  otu_table() %>%
  as.data.frame()

#Make the dataframe long for plotting
akks.long <- all_akks %>%
  gather(ASV, Abundance)

#Plot Akk ASV abundances
akks.boxplot <- akks.long %>%
  ggplot(aes(x = ASV, y = Abundance)) +
  geom_point() +
  geom_jitter()
akks.boxplot
```

It looks like taking more than just the one wouldn't contribute to anything. This implies that there is only one phylogroup present in this dataset. Taking the sequence and comparing to known sequences reveals that it is an AmI.
