---
title: "Ang Akkermansia ASVs"
author: "Katherine Mueller"
date: "7/8/2021"
output: html_document
---

#Load Libraries
```{r}
library(phyloseq)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

#Define Paths
```{r}
ang.rds <- "../../Data/Analysis/ang.rds"
out.dir <- "../../Data/Analysis"
```

#Load and Check Phyloseq Object
```{r}
ang.ps = read_rds(ang.rds)
print(ang.ps)
```

```{r}
sample_variables(ang.ps)
```

#Relative Abundance
```{r}
rel.ps <- ang.ps %>%
  transform_sample_counts(function(x) x/sum(x))
```


#Akkermansia ASV Identification
```{r}
akks.ps <- rel.ps %>%
  subset_taxa(Genus == "Akkermansia")
print(akks.ps)
```

Filter for Akk ASVs at a mean abundance of at least 0.01%
```{r}
mean_cutoff = 0.0001
mean_test = function(x) {
  mean(x) >= mean_cutoff
}

akks.0001mean.ps <- akks.ps %>%
  filter_taxa(mean_test, prune = TRUE)
print(akks.0001mean.ps)
```

```{r}
akks0001 <- akks.0001mean.ps %>%
  otu_table() %>%
  as.data.frame()
colnames(akks0001)
```

Double check that none of the other ASVs should be included
```{r}
#Get dataframe for akk ASV abundances
all_akks <- akks.ps %>%
  otu_table() %>%
  as.data.frame()

#Make the dataframe long for plotting
akks.long <- all_akks %>%
  gather(ASV, Abundance)

#Plot Akk ASV abundances
akks.boxplot <- akks.long %>%
  ggplot(aes(x = ASV, y = Abundance)) +
  geom_point() +
  geom_jitter()
akks.boxplot
```

It looks like taking more than just the one wouldn't contribute to anything. This implies that there is only one phylogroup present in this dataset. Taking the sequence and comparing to known sequences reveals that it is an AmI.

#Breakdown of Phylogroup Abundances
Even though I should only need to filter the Akkermansia ASVs to just the one, I still want to keep the output consistent with the other datasets being used.
```{r}
#Rename ASV sequences for ease of reading
colnames(akks0001)[1] <- "ASV1"
colnames(akks0001)
```

```{r}
#Determine presence/absence of each phylogroup
akks <- akks0001 %>%
  mutate(Akk_abundance = rowSums(.),
         has_AmI= ifelse(ASV1 >= 0.0001, "AmI_Positive", "AmI_Negative"),
         has_Akk = ifelse(Akk_abundance >= 0.0001, "Akk_Positive", "Akk_Negative"))

#Assign new variables
sample_data(rel.ps)$AmI_abund <- akks$ASV1
sample_data(rel.ps)$AmI_presence <- akks$has_AmI
sample_data(rel.ps)$Akk_abund <- akks$Akk_abundance
sample_data(rel.ps)$Akk_presence <- akks$has_Akk
```

```{r}
sample_variables(rel.ps)
```

Output the updated phyloseq object for later use
```{r}
saveRDS(rel.ps, file.path(out.dir, 'ang_wPhylogroups.rds'))
```

#Session Info
```{r}
sessionInfo()
```

