---
title: "Baumgartner Akkermansia ASVs"
author: "Katherine Mueller"
date: "10/14/2021"
output: html_document
---

#Load Libraries
```{r}
library(phyloseq)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

#Define Paths
```{r}
baumgartner.rds <- "../../Data/Analysis/baumgartner.rds"
out.dir <- "../../Data/Analysis"
```

Load and check phyloseq object
```{r}
baumgartner.ps = read_rds(baumgartner.rds)
print(baumgartner.ps)
```

```{r}
sample_variables(baumgartner.ps)
```

#Relative Abundance
```{r}
rel.ps <- baumgartner.ps %>%
  transform_sample_counts(function(x) x/sum(x))
```

#Akkermansia ASV Identification
```{r}
akks.ps <- rel.ps %>%
  subset_taxa(Genus == "Akkermansia")
print(akks.ps)
```

Filter for Akk ASVs at a mean abundance of at least 0.01%
```{r}
mean_cutoff = 0.0001
mean_test = function(x) {
  mean(x) >= mean_cutoff
}

akks.0001mean.ps <- akks.ps %>%
  filter_taxa(mean_test, prune = TRUE)
print(akks.0001mean.ps)
```

That didn't work.
Make a graph of all ASV abundances to see if maybe none of these ASVs pass the filter.
```{r}
#Get dataframe for akk ASV abundances
all_akks <- akks.ps %>%
  otu_table() %>%
  as.data.frame()

#Make the dataframe long for plotting
akks.long <- all_akks %>%
  gather(ASV, Abundance)

#Plot Akk ASV abundances
akks.boxplot <- akks.long %>%
  ggplot(aes(x = ASV, y = Abundance)) +
  geom_point() +
  geom_jitter()
akks.boxplot
```

It looks like at least three of these should pass the filter. Do the abundance values look ok or are we missing data?
```{r}
all_akks[[1]]
```

```{r}
mean(all_akks[[1]])
```

It looks like those NaN cells are messing things up. I'll replace them with 0.
```{r}
#Pull oTU table
akk.otus <- akks.ps %>%
  otu_table() %>%
  as.data.frame()

#Replace NaNs with 0s
good.akk.otus <- replace(akk.otus, is.na(akk.otus), 0)

#Construct fixed phyloseq object
sd = sample_data(akks.ps)
taxa = tax_table(akks.ps)

good.akks.ps <- phyloseq(otu_table(good.akk.otus, taxa_are_rows = FALSE), sd, taxa)
```

Check new akk phyloseq
```{r}
good.akks.ps
```

Let's try filtering again

Filter for Akk ASVs at a mean abundance of at least 0.01%
```{r}
mean_cutoff = 0.0001
mean_test = function(x) {
  mean(x) >= mean_cutoff
}

akks.0001mean.ps <- good.akks.ps %>%
  filter_taxa(mean_test, prune = TRUE)
print(akks.0001mean.ps)
```

```{r}
akks0001 <- akks.0001mean.ps %>%
  otu_table() %>%
  as.data.frame()
colnames(akks0001)
```

Taking the sequences of the four and comparing to known sequences reveals that ASV1 is an AmI, ASV2 is an AmI, ASV3 is an AmII, ASV4 is an AmI, ASV5 is an AmIV, and ASV6 is an AmI.

#Breakdown of Phylogroup Abundances
```{r}
#Rename ASV sequences for ease of reading
colnames(akks0001)[1] <- "ASV1"
colnames(akks0001)[2] <- "ASV2"
colnames(akks0001)[3] <- "ASV3"
colnames(akks0001)[4] <- "ASV4"
colnames(akks0001)[5] <- "ASV5"
colnames(akks0001)[6] <- "ASV6"
colnames(akks0001)
```

```{r}
#Determine presence/absunce of each phylogroup
akks <- akks0001 %>%
  mutate(Akk_abundance = rowSums(.),
         has_AmI = ifelse(ASV1 + ASV2 + ASV4 + ASV6 >= 0.0001, "AmI_Positive", "AmI_Negative"),
         has_AmII = ifelse(ASV3 >= 0.0001, "AmII_Positive", "AmII_Negative"),
         has_AmIV = ifelse(ASV5 >= 0.0001, "AmIV_Positive", "AmIV_Negative"),
         has_Akk = ifelse(Akk_abundance >= 0.0001, "Akk_Positive", "Akk_Negative"))

#Assign new variables
sample_data(rel.ps)$AmI_abund <- (akks$ASV1 + akks$ASV2 + akks$ASV4 + akks$ASV6)
sample_data(rel.ps)$AmI_presence <- akks$has_AmI
sample_data(rel.ps)$AmII_abund <- akks$ASV3
sample_data(rel.ps)$AmII_presence <- akks$has_AmII
sample_data(rel.ps)$AmIV_abund <- akks$ASV5
sample_data(rel.ps)$AmIV_presence <- akks$has_AmIV
sample_data(rel.ps)$Akk_abund <- akks$Akk_abundance
sample_data(rel.ps)$Akk_presence <- akks$has_Akk
```

```{r}
sample_variables(rel.ps)
```

Output the updated phyloseq object for later use
```{r}
saveRDS(rel.ps, file.path(out.dir, 'baumgartner_wPhylogroups.rds'))
```

#Session Info
```{r}
sessionInfo()
```
